---
# title: "Analyze behavioural measures of specificity experiment"
# author: "Ryan Law"
# date: "`r format(Sys.time(), "%B %d, %Y")`"
output:
  html_document: default
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
library(effects)
library(ggplot2)
```

# Analyses of behavioural data from MEEG study of conceptual specificity in words and phrases

Task: word and phrase comprehension (yes/no questions on 10% of trials)
Stimulus presentation: RSVP

```{r read in logfiles}

# read in log files from all subjects
results_dir <- "U:/ownCloud/projects/NocturnalBird/results/behavioural"
list_of_logfiles <- list.files(path = 'U:/ownCloud/projects/NocturnalBird/scripts/analysis/behavioural/data',
                               recursive = TRUE,
                               pattern = "\\.csv$",
                               full.names = TRUE)
group_data <- readr::read_csv(list_of_logfiles, id = "file_name")

# select relevant columns from log files
group_data = select(group_data, participant, block_nr, trial_nr, item_nr, word1, word2, probe, response, RT, hit, composition, specificity, zipf_noun) 

# select specificity trials and filter out filler trials
group_data <- subset(group_data, !is.na(group_data$specificity))

# select trials with a task question probe (only 10% trials have a probe)
group_data <- subset(group_data, !is.na(group_data$probe))

# participant 25 flipped their button presses so need to flip their hits
group_data <- group_data %>%
  mutate(
    hit = ifelse(participant == 25, 1 - hit, hit)
  )

# transform data and set column format
group_data = mutate(group_data,
                   RT = log10(RT), # log10-transform RT
                   participant = factor(participant),
                   item_nr = factor(item_nr),
                   specificity = factor(specificity))
group_data = as_tibble(group_data)
group_data
```

```{r accuracy descriptives}

# get accuracy (mean of hit per condition per participant)
accuracy <- group_data %>%
  group_by(participant, specificity) %>%
  summarize(
    accuracy = mean(hit),
    .groups = "drop"
  )

# get some descriptives
summary_accuracy_by_specificity <- accuracy %>%
  group_by(specificity) %>%
  summarize(
    mean_accuracy = mean(accuracy),
    sd_accuracy = sd(accuracy),    # Standard deviation for each specificity level
    .groups = "drop"
  )
print(summary_accuracy_by_specificity)

summary_accuracy_by_participant <- accuracy %>%
  group_by(participant) %>%
  summarize(
    mean_accuracy = mean(accuracy),
    sd_accuracy = sd(accuracy),    # Standard deviation for each specificity level
    .groups = "drop"
  )
print(summary_accuracy_by_participant)

# plot accuracy by participant
fig_acc.ptp <- ggplot(accuracy, aes(x = as.factor(participant), y = accuracy)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 2) +
  theme_minimal() +
  labs(
    title = "Accuracy by Participants with Specificity Variability",
    x = "Participant",
    y = "Accuracy"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate participant labels for clarity
fig_acc.ptp
ggsave(filename = file.path(results_dir, "fig_accuracy_by_participant.jpg"), 
       plot = fig_acc.ptp, width = 8, height = 6, dpi = 300)

# plot accuracy by specificity
fig_acc.spec <- ggplot(accuracy, aes(x = as.factor(specificity), y = accuracy)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 2) +
  theme_minimal() +
  labs(
    title = "Accuracy by Participants with Specificity Variability",
    x = "Participant",
    y = "Accuracy"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate participant labels for clarity
fig_acc.spec
ggsave(filename = file.path(results_dir, "fig_accuracy_by_specificity.jpg"), 
       plot = fig_acc.spec, width = 4, height = 6, dpi = 300)


```
Fitting linear mixed effects models. Models that are more complex than these do not converge.

```{r lmer_acc }

# linear mixed model with by-participant and by-item varying intercepts
model_acc.base = glmer(hit ~ zipf_noun + (1|participant) + (1|item_nr), data=group_data, family='binomial')
model_acc.full = glmer(hit ~ zipf_noun + specificity + (1|participant) + (1|item_nr), data=group_data, family='binomial')

print(summary(model_acc.base))
print(summary(model_acc.full))
anova(model_acc.base, model_acc.full)
print(Anova(model_acc.full))

```

From the `lmer` model, it looks like there is a significant effect of zipf frequency of the noun and no effect of specificity. It could be that our task is not sensitive to specificity the way picture matching tasks might (cf. Rogers & McClelland, 2003). 

```{r plot relationship between zipf and hit rate}
predicted <- data.frame(
  zipf_noun = group_data$zipf_noun,
  accuracy = predict(model_acc.full, re.form = NA)
)
ggplot(predicted, aes(x = zipf_noun, y = accuracy)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r plot emmeans results Accuracy}

# # Back-transform to the response scale
# emmeans_result_response_scale <- regrid(model.emm)
# summary(emmeans_result_response_scale)
# 
# # Plot the estimated marginal means
# emm_plot <- emmip(model.emm, specificity, CIs = TRUE) + 
#             theme_minimal()
# ggsave(filename = file.path(results_dir, "fig_ACC_emmeans.jpg"), 
#        plot = emm_plot, width = 8, height = 6, dpi = 300)
# Extract means and confidence intervals for plotting
# emm_means <- as.data.frame(model.emm)

# # Add the pairwise comparisons to the plot
# comparison_labels <- data.frame(
#   concreteness = rep(levels(group_data$concreteness), each = length(levels(group_data$denotation))),
#   denotation = rep(levels(group_data$denotation), times = length(levels(group_data$concreteness))),
#   significance = comparison_results.df$significance
# )
# 
# # Add text for significance
# emm_plot + geom_text(data = comparison_labels, aes(x = factor2, y = estimate, label = significance),
#                      vjust = -0.5, color = "red")


```

For reaction times, we similarly plot the distribution of reaction times by specificity levels.

```{r group level RT}
group_data = group_data[group_data$hit == 1, ] # only analysing reaction times from trials answered correctly

# make density plot of RT distribution
plot <- group_data %>% ggplot(aes(x = RT, fill = specificity)) +
 geom_density(alpha = 0.5) + theme_minimal() +
 scale_fill_brewer(palette = 'PuOr')
ggsave(filename = file.path(results_dir, "fig_rt_by_specificity_density.jpg"), 
       plot = plot, width = 8, height = 6, dpi = 300)

# make summary stats for RT for each condition
rt.spec <- group_data %>%
  group_by(specificity) %>%
  summarise(
    count = n(),
    mean = mean(RT),
    sd = sd(RT),
    se = sd(RT) / sqrt(n()),
    median = median(RT),
    min = min(RT),
    max = max(RT)
  )
print(rt.spec)

fig_rt.spec <- ggplot(rt.spec, aes(x = factor(specificity), y = mean)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  theme_minimal() +
  labs(
    title = "Mean Reaction Time by specificity",
    x = "Specificity",
    y = "Mean Reaction Time (s)"
  )
fig_rt.spec
ggsave(filename = file.path(results_dir, 'fig_rt_by_specificity.jpg'), 
       plot = fig_rt.spec, width = 8, height = 6, dpi = 300)

# write.csv(summary_RT_specificity, file = file.path(results_dir, "summary_stats_RT_specificity.csv"), row.names = FALSE)

```

```{r lmer_rt }

# fit linear mixed models
# by-participant and by-item varying intercepts
model_rt.base = lmer(RT ~ zipf_noun + (1|participant) + (1|item_nr), data=group_data, REML=TRUE)
model_rt.full = lmer(RT ~ zipf_noun * specificity + (1|participant) + (1|item_nr), data=group_data, REML=TRUE)
print(summary(model_rt.base))
print(summary(model_rt.full))
print(anova(model_rt.base, model_rt.full))
print(Anova(model_rt.full))
# # estimate marginal means
# model.emm <- emmeans(model_rt, ~ specificity)
# 
# # compute pairwise comparisons
# comparison_results <- pairs(model.emm, adjust='tukey')
# # add asterisks for significant values
# comparison_results.df <- as.data.frame(comparison_results)
# comparison_results.df$significance <- cut(comparison_results.df$p.value, 
#                                     breaks = c(-Inf, 0.001, 0.01, 0.05, Inf), 
#                                     labels = c("***", "**", "*", ""))
# print(comparison_results.df)
```

Looks like there is also no effect of specificity on reaction times.

```{r plot relationship between zipf and reaction time}
# predicted <- data.frame(
#   zipf_noun = group_data$zipf_noun,
#   rt = predict(model_rt.full, re.form = NA)
# )
# ggplot(predicted, aes(x = zipf_noun, y = rt)) +
#   geom_point() +
#   geom_smooth(method = "lm")
```

```{r plot emmeans results RT}

# # Back-transform to the response scale
# emmeans_result_response_scale <- regrid(model.emm)
# summary(emmeans_result_response_scale)
# 
# # Obtain the overall EMM for the 'concreteness' factor
# emm_concreteness <- emmeans(model.emm, ~ specificity)
# summary(emm_concreteness) # there is no interaction so safe to interpret these
# 
# # Plot the estimated marginal means
# emm_plot <- emmip(model.emm, specificity, CIs = TRUE) + 
#             theme_minimal()
# ggsave(filename = file.path(results_dir, "fig_RT_emmeans.jpg"), 
#        plot = emm_plot, width = 8, height = 6, dpi = 300)
# Extract means and confidence intervals for plotting
# emm_means <- as.data.frame(model.emm)

# # Add the pairwise comparisons to the plot
# comparison_labels <- data.frame(
#   concreteness = rep(levels(group_data$concreteness), each = length(levels(group_data$denotation))),
#   denotation = rep(levels(group_data$denotation), times = length(levels(group_data$concreteness))),
#   significance = comparison_results.df$significance
# )
# 
# # Add text for significance
# emm_plot + geom_text(data = comparison_labels, aes(x = factor2, y = estimate, label = significance),
#                      vjust = -0.5, color = "red")


```